### 2022-02-18 sections/simulated_samples.tex

%-------------------------------------------------------------------------------
\section{Signal and background modelling}
\label{sec:signal_background_model}
%-------------------------------------------------------------------------------

Signal and most background processes are modelled using Monte Carlo (MC) simulation.
After the event preselection, the main background is $\ttbar$ production, often in association with jets, denoted by $\ttbar$+jets in the following.
Small contributions arise from single-top-quark, $W/Z$+jets, multijet and diboson ($WW,WZ,ZZ$) production, as well as from the associated 
production of a vector boson $V$ ($V=W,Z$) or a Higgs boson and a $\ttbar$ pair ($\ttbar V$ and $\ttbar H$). All backgrounds 
with prompt leptons, i.e.\ those originating from the decay of a $W$ boson, a $Z$ boson, or a $\tau$-lepton,
are estimated using samples of simulated events and initially normalised to their theoretical cross sections.
In the simulation, the top-quark and SM Higgs boson masses are set to $172.5~\gev$ and $125~\gev$, respectively,
and the Higgs boson is forced to decay into $H\to \tau\tau$ with branching ratio calculated using \textsc{Hdecay}~\cite{Djouadi:1997yw}.  
Backgrounds with non-prompt electrons or muons, with photons or jets misidentified as electrons, or with jets misidentified as $\had$ candidates, 
generically referred to as fake leptons, are estimated using data-driven methods. 
The background prediction is further improved during the statistical analysis by performing a likelihood 
fit to data using several signal-depleted analysis regions, as discussed in Sections~\ref{sec:strategy_Htautau}.

%-------------------------------------------------------------------------------
\subsection{Simulated signal and background processes}
\label{sec:simulations}
%-------------------------------------------------------------------------------

@@

Samples of simulated $\ttbar \to WbHq$ events were generated with the next-to-leading-order (NLO) generator\footnote{In the following, 
the order of a generator should be understood as referring to the order in the strong coupling constant at which the matrix-element calculation 
is performed.} {\powheg}~v2 \cite{Frixione:2007nw,Nason:2004rx,Frixione:2007vw,Alioli:2010xd}
%{\amcatnlolong}~2.4.3~\cite{Alwall:2014hca}  (referred to in the following as {\amcatnlo})
with the NNPDF3.0 NLO~\cite{Ball:2014uwa} parton distribution function (PDF) set and interfaced to {\pythia} 8.212~\cite{Sjostrand:2007gs} with the NNPDF2.3 LO~\cite{Ball:2012cx} PDF set for the modelling of parton showering, hadronisation, and the underlying event. 
The A14~\cite{ATLASUETune4} set of tuned parameters in {\pythia} controlling the description of multiparton interactions and  
initial- and final-state radiation, referred to as the tune, was used.
The signal sample is normalised to the same total cross section as used for the inclusive $t\bar{t}\to WbWb$ sample (see discussion below) and
assuming a benchmark branching ratio of $\BR_{\mathrm{ref}}(t\to qH)=0.1\%$.
The case of both top quarks decaying into $qH$ is neglected in the analysis given the existing upper limits on $\BR(t \to qH)$ (Section~\ref{sec:intro}).

@@

@@

\begin{table}
\footnotesize
\centering
\caption{Overview of the MC generators used for the main signal and background samples}
\begin{tabular}[h]{l|c|c|c|c|c|c}
\hline \hline
\multirow{2}{*}{Process} & \multicolumn{2}{c|}{Generator} & \multicolumn{2}{c|}{PDF set} & \multirow{2}{*}{Tune} & \multirow{2}{*}{Order} \\ \cline{2-5}
        &  ME   &  PS    &  ME  & PS &   &  \\\hline
\textit{TT} Signal & Powheg & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
\textit{ST} Signal & MadGraph5\_aMC@NLO & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
$W/Z$+jets & \multicolumn{2}{c|}{Sherpa 2.2.1} & \multicolumn{2}{c|}{NNPDF30NNLO} & Sherpa & NLO/LO \\ \hline
\ttbar & Powheg & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
Single top & Powheg & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
$t\bar{t}X$ & MadGraph5\_aMC@NLO & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
Diboson & \multicolumn{2}{c|}{Sherpa 2.2.1} & \multicolumn{2}{c|}{NNPDF30NNLO} & Sherpa & NLO/LO \\ \hline\hline
\end{tabular}
\label{mob}
\end{table}

@@

\subsubsection{Fake electrons and muons}
Electrons and muons only appear in the leptonic channels where there is exactly one electron or muon and at least one $\had$. Since the fake rate of $\had$ candidate is much larger than electron or muon, in the majority of the fake lepton events, the $\had$ candidate is misidentified and the electron or muon is real, which is also proved by monte carlo studies. However the multi-jet background has large event rate and can contribute with both electron or muon and $\had$ faked, especially to $t_l\had$ where the fake-lepton dominates.
%and the jet multiplicity is low.
This kind of event is modelled by data-driven ABCD method by cutting on $\met$ and the tight lepton isolation using PLIV, which is widely used in ATLAS physics analyses.

@@

\subsubsection{Fake $\tau$-lepton candidates}
\label{sec:faketaus}
The background with one or more fake $\had$ candidates mainly arises from $\ttbar$ or
multijet production, depending on the search channels.
Studies based on the simulation show that, for all the above processes, fake $\had$ candidates primarily result from the
misidentification of light-quark jets, with the contribution from $b$-quarks and gluon jets playing a subdominant role.
It is also found that the fake rate decreases for all jet flavours as the $\had$ candidate $\pt$ increases.

In the leptonic channels, the events with real electron or muon and fake taus are modelled by calibrating monte carlo with scale factors
derived from the dedicated $t\bar t$
control regions ($CR_{tt}$) using the SM $t\bar t$ decay of dilepton events and semileptonically double-btagged lepton-jet events, summarized in Table~\ref{tab:srcr},
aimed at fake taus from different origins. There are four kinds of fake taus that need to be calibrated: Type1) the fake taus from $W$-jets ($\tau_{W}$)
with the opposite charge to the lepton;
Type2) $\tau_{W}$'s with the same charge as the lepton; Type 3) the fake taus from $b$-jets; Type4) the fake taus from other origins (mainly radiations).
These control regions are similar to the signal regions but with an addition $b$-jet or lepton defined as following:
$t_lt_l1b\thad$, $t_lt_l2b\thad$, $t_lt_h2b\thad$-2jSS, $t_lt_h2b\thad$-2jOS, $t_lt_h2b\thad$-3jSS, and $t_lt_h2b\thad$-3jOS.
The di-lepton regions ($t_lt_l1b\thad$ and $t_lt_l2b\thad$) are used to calibrate the Type3 and Type4 fake taus. The regions of
$t_lt_h2b\thad$-2jOS and $t_lt_h2b\thad$-3jOS are used to calibrate Type1 fake taus.
Similarly for the Type2, the regions of $t_lt_h2b\thad$-2jSS and $t_lt_h2b\thad$-3jSS are used.
A simultaneous fit to data is made to derive the scale factors for the fake taus in MC, which consist of total 24 parameters
including four types, 3 $\pt$ bins (25-35, 35-45, 45- GeV), 2 bins for 1 and 3 prong taus separately.
The values of these scale factors range from 0.3 to 1.28 and are applied to the corresponding signal regions with single b-tagged jet.

In the $t_l\thadhad$ channel, both taus can be fake, so the calibration is applied to each tau separately, following the same procedure as $\tlhad$ channels using the
lepton and fake tau charges, then the scale factors are multiplied together. The nominal value of the scale factors will vary along with other uncertainties from
combined performance (CP) recommendations and theory uncertainties in the final fit.
%In the control regions with single lepton, $\met > 20$GeV and at least 2 light jets and 2 b-tagged jets are required to ensure that QCD contribution is negligible.
%The calibration is done depending on different source of the fake taus and \pt. The calibration factor are derived in the dedicated
%control regions discussed in Section~\ref{sec:strategy_Htautau}.

In the hadronic channels, this background is estimated partially from data by defining control regions (CR) enriched in fake $\had$
candidates via loosened $\had$ requirements with proper fake factors and partially from monte carlo~\cite{ATLAS-CONF-2021-044}.
These CRs do not overlap with the main search regions (SRs), discussed in Section~\ref{sec:strategy_Htautau}.
The CR selection requirements are analogous to those used to define the different SRs, except
that the trailing $\had$ candidate is required to fail the medium $\had$ identification.
The events with fake trailing $\had$ candidate can be calculated by rescaling the templates of fake taus in the Fake-CR with proper fake factors (FF).
The templates are acquired by subtracting all MC background contributions with real sub-leading taus from data. The FFs are computed as
the ratio of the Data-MC($\mathrm{real~tau}$) yields passing the medium tau ID to which failing the medium tau ID from different control regions and
the differences are treated as part of systematics.

>>>>

%-------------------------------------------------------------------------------
%-------------------------------------------------------------------------------
\section{Simulated samples}
\label{sec:simulations}
%-------------------------------------------------------------------------------

@@

Samples of simulated $\ttbar \to WbHq$ ($tt(qH)$) events were generated with the next-to-leading-order (NLO) generator\footnote{In the following, 
the order of a generator should be understood as referring to the order in the strong coupling constant at which the matrix-element (ME) calculation 
is performed.} {\powheg}~v2 \cite{Frixione:2007nw,Nason:2004rx,Frixione:2007vw,Alioli:2010xd}
%{\amcatnlolong}~2.4.3~\cite{Alwall:2014hca}  (referred to in the following as {\amcatnlo})
with the NNPDF3.0 NLO~\cite{Ball:2014uwa} parton distribution function (PDF) set and interfaced to {\pythia} 8.212~\cite{Sjostrand:2007gs} with the NNPDF2.3 LO~\cite{Ball:2012cx} PDF set for the modelling of parton showering (PS), hadronisation, and the underlying event. 
The A14~\cite{ATLASUETune4} set of tuned parameters in {\pythia} controlling the description of multiparton interactions and  
initial- and final-state radiation, referred to as the tune, was used.
The signal sample is normalised to the same total cross section as used for the inclusive $t\bar{t}\to WbWb$ sample (see discussion below) and
assuming a benchmark branching ratio of $\BR_{\mathrm{ref}}(t\to qH)=0.1\%$.
The case of both top quarks decaying into $qH$ is neglected in the analysis given the existing upper limits on $\BR(t \to qH)$ (Section~\ref{sec:intro}).

@@

Samples of $WH$ and $ZH$, collectively referred to as $VH$ are generated using {\powheg}~v2 \cite{Frixione:2007nw,Nason:2004rx,Frixione:2007vw,Alioli:2010xd}
and interfaced to {\pythia}~8.210 with the NNPDF2.3 LO PDF set and the A14 tune.
The contribution of $tH$ associated production is also considered as part of SM Higgs background.
The sample is generated using {\amcatnlolong}~2.6.2~\cite{Alwall:2014hca} and interfaced with {\pythia}~8.210 with CT10PDF
and A14 tune. The MC predictions are normalised using the NLO cross section recommended in Ref.~\cite{deFlorian:2016spz}.

@@

\begin{table}
\footnotesize
\centering
\caption{Overview of the MC generators used for the main signal and background samples}
\begin{tabular}[h]{l|c|c|c|c|c|c}
\hline \hline
\multirow{2}{*}{Process} & \multicolumn{2}{c|}{Generator} & \multicolumn{2}{c|}{PDF set} & \multirow{2}{*}{Tune} & \multirow{2}{*}{Order} \\ \cline{2-5}
        &  ME   &  PS    &  ME  & PS &   &  \\\hline
$tt(qH)$ Signal & Powheg & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
$tH$ Signal & MadGraph5\_aMC@NLO & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
$W/Z$+jets & \multicolumn{2}{c|}{Sherpa 2.2.1} & \multicolumn{2}{c|}{NNPDF30NNLO} & Sherpa & NLO/LO \\ \hline
\ttbar & Powheg & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
Single top & Powheg & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
$t\bar{t}X$ & MadGraph5\_aMC@NLO & Pythia8 & NNPDF30NLO & NNPDF23LO & A14 & NLO \\ \hline
Diboson & \multicolumn{2}{c|}{Sherpa 2.2.1} & \multicolumn{2}{c|}{NNPDF30NNLO} & Sherpa & NLO/LO \\ \hline\hline
\end{tabular}
\label{mob}
\end{table}

@@

\subsubsection{Fake electrons and muons}
Electrons and muons only appear in the leptonic channels where there is exactly one electron or muon and at least one $\had$. Since the fake rate of $\had$ candidate is much larger than for electron or muon, in the majority of the fake lepton events, the $\had$ candidate is misidentified and the electron or muon is prompt based on MC studies. However the multi-jet background has large event rate and can contribute with both electron or muon and $\had$ faked, especially to $t_l\had$ where the fake-lepton dominates.
%and the jet multiplicity is low.
The contribution from these events is estimated with a data-driven method called ABCD.
The estimate is based on the number of tight and loose leptons that passed and failed the PLIV cut in the low and high $\met$ regions. The number of
fake leptons in the signal region is evaluated by rescaling the number of loose leptons in the high $\met$ region with the ratio of tight and loose leptons in the
low $\met$ region. The contribution of prompt leptons in the number of loose leptons is subtracted using MC simulation.

%This kind of event is modelled by data-driven ABCD method
%by cutting on $\met$ and the tight lepton isolation using PLIV, which is widely used in ATLAS physics analyses.

@@

### 2022-02-18 sections/strategy_Htautau.tex

%%%%%%%%%%%%%%%
\begin{table*}[t!]
  \caption{\small{Discriminating variables used in the training of the BDT for hadronic channel.
      The values in percent (\%) represent the separation and importance of each variable.
    The descriptions of each variable are provided in the following text.}}
\label{tab:importance_xTFW}
\input{tables/xTFW_Merged}
\end{table*}

\begin{table*}[t!]
  \caption{\small{Discriminating variables used in the training of the BDT for leptonic channel.
      The values in percent (\%) represent the separation and importance of each variable.
    The descriptions of each variable are provided in the following text.}}
\label{tab:importance_tthML}
\input{tables/tthML_Merged}
\end{table*}

@@

\caption{The leading $\thad$ $p_T$  distributions are compared between the expected background and $tuH$ signals in: $t_l\thadhad$ (a1),  $t_l\tauhad$-1j (a2),  $t_l\tauhad$-2j (a3), $t_h\tlhad$-2j (b1), $t_h\tlhad$-3j (b2), $t_h\thadhad$-2j (b3), and $t_h\thadhad$-3j (c1) before the fit to data ('Pre-Fit'). The uncertainty band includes both the statistical and systematic uncertainties in the background prediction. Underflow and overflow bins are included respectively in the first and last bins. Rare including single top, and V+jets and only subtau real refers to $t\bar{t}$ MC with leading tau faked by other object, subleading tau from real contribution. The lower panels show the ratios of the data ('Data') to the background prediction ('Bkg.')}

>>>>

%%%%%%%%%%%%%%%
\begin{table*}[t!]
\caption{\small{Discriminating variables used in the training of the BDT. The values represent the importance rankings of each variable and cross($\times$) denotes the variable is not used in the training. The descriptions of each variable are provided in the following text.}}
\label{tab:importance}
 \begin{tabular}{cccccccc} \toprule\toprule
& $t_{l}\tauhad$-1j                                  &  $t_{h}\tlhad$-2j   &  $t_{l}\tauhad$-2j & $t_{h}\tlhad$-3j & $t_l\thadhad$     & $t_h\thadhad$-2j & $t_h\thadhad$-3j       \\\midrule 
 $m_{\text{W}}$                                      & $\times$  &  $\times$           &  $\times$         & $9$      & $\times$      & $6$      & $7$\\
 $\chi^{2}$                                          & $\times$  &  $\times$           &  $\times$         & $14$     & $\times$      & $\times$ & $\times$      \\
 \text{max}($\eta_{\tau}$)                           & $4$       &  $\times$           &  $4$              & $\times$ & $10$          & $\times$ & $\times$       \\
 $m^{\text{T}}_{\text{W}}$                           & $10$      &  $\times$           &  $8$              & $\times$ & $13$          & $\times$ & $\times$        \\
 $m_{\tau,\tau}$                                     & $\times$  &  $2$                &  $\times$         & $3$      & $\times$      & $1$      & $1$          \\
 $m_{\text{t},\text{SM}}$                            & $\times$  &  $1$                &  $\times$         & $2$      & $\times$      & $3$      & $4$          \\
 $p_{\text{T},\ell}$                                 & $12$      &  $15$               &  $12$             & $17$     & $\times$      & $\times$ & $\times$        \\
 $m_{\text{t},\text{FCNC}}$                          &  $\times$ &  $\times$           &  $\times$         & $\times$ & $\times$      & $10$     & $6$\\
 $m_{\text{t},\text{SM},\text{vis}}$                 & $3$       &  $\times$           &  $5$              & $\times$ & $4$           & $\times$ & $\times$        \\
 $p_{\text{T},\tau}$                                 & $1$       &  $4$                &  $1$              & $1$      & $5$           & $11$   & $10$           \\
 $\met$                                              & $5$       &  $11$               &  $10$             & $13$     & $6$           & $7$    & $13$          \\
 $m_{\tau\tau,\text{vis}}$                           & $10$      &  $14$               &  $11$             & $6$      & $1$           & $2$    & $2$          \\
 $E_{\text{vis}~\tau,1}/E_{\tau,1}$                  & $\times$  &  $10$               &  $\times$         & $12$     & $\times$      & $8$    & $8$          \\
 $E_{\text{vis}~\tau,2}/E_{\tau,2}$                  & $\times$  &  $7$                &  $\times$         & $4$      & $\times$      & $9$    & $11$         \\
 $P_{\text{T,\tauhadvis}} $                          & $\times$  &  $\times$           &  $\times$         & $\times$ & $9$           & $\times$ & $\times$        \\
 $m_{\text{t},\text{FCNC},\text{vis}}$               & $\times$  &  $\times$           &  $\times$         & $\times$ & $3$           & $\times$ & $\times$       \\
 $\Delta\phi(\tau\tau,\met)$                         & $\times$  &  $6$    			   &  $\times$         & $16$     & $\times$      & $13$   & $12$         \\
 $\met\text{centrality}$                             & $\times$  &  $13$               &  $\times$         & $15$     & $\times$      & $12$   & $9$         \\
 \text{Min}($m_{\tau\tau \text{q-jet}}$)             & $9$       &  $\times$           &  $3$              & $\times$ & $14$          & $\times$ & $\times$        \\
 $\Delta R(\ell,\tau)$                               & $8$       &  $9$                &  $9$              & $10$     & $15$          & $\times$ & $\times$        \\
 $\Delta R(\tau,\tau)$                               & $\times$  &  $\times$           &  $\times$         & $\times$ & $2$           & $4$    & $3$             \\
 $\Delta R(\ell,\text{b-jet})$                       & $2$       &  $3$                &  $2$              & $8$      & $12$          & $\times$ & $\times$        \\
 $\Delta R(\tau,\text{b-jet})$                       & $6$       &  $5$                &  $6$              & $7$      & $11$          & $\times$ & $\times$       \\
 $\Delta R(\ell+\text{b-jet},\tau\tau )$             & $\times$  &  $\times$           &  $\times$         & $\times$ & $7$           & $\times$ & $\times$        \\
 $\Delta R(\tau,\text{light-jet})$                   & $7$       &  $8$                &  $7$              & $5$      & $8$           & $5$    & $5$    \\
 \text{Min}($m_{\text{light-jet},\text{light-jet}}$) & $\times$  &  $12$               &  $\times$         & $11$     & $\times$      & $\times$ & $\times$        \\
 \bottomrule\bottomrule\\
\end{table*}

%%%%%%%%%%%%%%%
%%%  \begin{table*}[t!]
%%%    \caption{\small{Discriminating variables used in the training of the BDT for hadronic channel.
%%%        The values in percent (\%) represent the separation and importance of each variable.
%%%      The descriptions of each variable are provided in the following text.}}
%%%  \label{tab:importance_xTFW}
%%%  \input{tables/xTFW_Merged}
%%%  \end{table*}
%%%  
%%%  \begin{table*}[t!]
%%%    \caption{\small{Discriminating variables used in the training of the BDT for leptonic channel.
%%%        The values in percent (\%) represent the separation and importance of each variable.
%%%      The descriptions of each variable are provided in the following text.}}
%%%  \label{tab:importance_tthML}
%%%  \input{tables/tthML_Merged}
%%%  \end{table*}

@@

\caption{The leading $\thad$ $p_T$  distributions are compared between the expected background and $tuH$ signals in: $t_l\thadhad$ (a1),  $t_l\tauhad$-1j (a2),  $t_l\tauhad$-2j (a3), $t_h\tlhad$-2j (b1), $t_h\tlhad$-3j (b2), $t_h\thadhad$-2j (b3), and $t_h\thadhad$-3j (c1) before the fit to data ('Pre-Fit'). The uncertainty band includes both the statistical and systematic uncertainties in the background prediction. Underflow and overflow bins are included respectively in the first and last bins. Rare including single top, and $V$+jets and only subtau real refers to $t\bar{t}$ MC with leading tau faked by other object, subleading tau from real contribution. The lower panels show the ratios of the data to the background prediction.}
